### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c75934d4c3d789d4d02b04bd3adedf4ed3efeaa9 🚀

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{38,39} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with more
than 1,400 estimated species.\textsuperscript{40,41} They exhibit a
broad variety of habitat use, behaviour, and feeding strategies, putting
them at key positions in the delivery and provisioning of several
ecosystem services, tied to important ecosystem-derived benefits to
humans.\textsuperscript{42} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{43,44} and
vectors of pathogens that put a risk on human
health;\textsuperscript{45,46} some other species are essential links in
many seed-dispersal networks.\textsuperscript{47} However, many of these
species face a high risk of extinction, particularly given persecution
and killings that sometimes follows from messaging about their role in
disease emergence. Areas where bats, viruses, and humans co-occur are
not always hotspots of risk for human heath; as such, developing more
precise ways to map zoonotic hazards can help bats and humans coexist
safely, and support the conservation of these important and unique
animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{19} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{48}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,49} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{26,50,51} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{52} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both (add cites for DOIs:
10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{53} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{54} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{55} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{56} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{57} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{58} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{59} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{60} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{61} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{62} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{63} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{64} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{65} v1.6.12
with ModelFinder\textsuperscript{66} ultrafast bootstrap
approximation\textsuperscript{67} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{68} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{69,70} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}

\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
virome diversity in bats. \emph{Current Opinion in Virology}
\textbf{52}, 192--202 (2022).}

\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
synthesis of ecological and evolutionary determinants of bat diversity
across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}

\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
World: A taxonomic and geographic database. \url{https://batnames.org/}
(2020).}

\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
\textbf{2013}, e187415 (2013).}

\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
\CSLLeftMargin{43. }
\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
a Changing World}. (Springer International Publishing, 2016).
doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}

\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
\textbf{320}, 70--70 (2008).}

\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
\emph{PLOS ONE} \textbf{8}, e77183 (2013).}

\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
Research} \textbf{40}, 10--24 (2013).}

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
\emph{PLOS ONE} \textbf{6}, e17395 (2011).}

\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
sarbecoviruses across China for the origin tracing of SARS-CoV and
SARS-CoV-2}. (2021)
doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{49. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
relationship between large-scale environmental determinants and bat
species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
(2018).}

\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
effect of local land use and loss of forests on bats and nocturnal
insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{53. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
mammal tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
(2019).}

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
variance of community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
of beta diversity in the Barro Colorado Island forest dynamics plot,
Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}

>>>>

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{38,39} They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to humans.\textsuperscript{40} Over
two-thirds of bats are know to be either obligate or facultative
insectivores, therefore actively contributing for agricultural pest
control,\textsuperscript{41,42} and vectors of pathogens that put a risk
on human health;\textsuperscript{43,44} some other species are essential
links in many seed-dispersal networks.\textsuperscript{45} However, many
of these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{19} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{46}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,47} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{26,48,49} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{50} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both (add cites for DOIs:
10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{51} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{52} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{53} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{54} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{55} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{56} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{57} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{58} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{59} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{60} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{61} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{62} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{63} v1.6.12
with ModelFinder\textsuperscript{64} ultrafast bootstrap
approximation\textsuperscript{65} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{66} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{67,68} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{69} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
synthesis of ecological and evolutionary determinants of bat diversity
across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}

\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
World: A taxonomic and geographic database. \url{https://batnames.org/}
(2020).}

\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
\textbf{2013}, e187415 (2013).}

\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
\CSLLeftMargin{41. }
\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
a Changing World}. (Springer International Publishing, 2016).
doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}

\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
\textbf{320}, 70--70 (2008).}

\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
\emph{PLOS ONE} \textbf{8}, e77183 (2013).}

\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
Research} \textbf{40}, 10--24 (2013).}

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
\emph{PLOS ONE} \textbf{6}, e17395 (2011).}

\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
sarbecoviruses across China for the origin tracing of SARS-CoV and
SARS-CoV-2}. (2021)
doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
relationship between large-scale environmental determinants and bat
species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
(2018).}

\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
\CSLLeftMargin{49. }
\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
effect of local land use and loss of forests on bats and nocturnal
insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{51. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
mammal tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
(2019).}

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
variance of community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
of beta diversity in the Barro Colorado Island forest dynamics plot,
Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}

\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
ecological uniqueness over broad spatial extents using species
distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
Predicting the global mammalian viral sharing network using
phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}

\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
geography of lake carbon cycling. \emph{Limnology and Oceanography
Letters} \textbf{3}, 49--56 (2018).}
